---
title: "Airbnb - Regression Model"
author: "Othman Bensouda Koraichi"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```

```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(kableExtra)
library(corrplot)
library(ggfortify)
library(huxtable)
library(leaflet)
library(reshape2)
```

# AirBnB analytics - Beijing

## Introduction 

The following project focuses on the analysis of an [AirBnB dataset](http://data.insideairbnb.com/china/beijing/beijing/2020-06-19/data/listings.csv.gz). A data cleaning has been firstly conducted to ensure reliable data. The output of the is a multivariate regression model that predicts the total cost of two people to stay at an AirBnB property in Beijing for 4 nights.  

## Executive summary

The dataset obtained from airBnB contained a vast amount of information which allows for the development of a broad range of insights into the operations of airBnB in Beijing. My group and I eliminated over half the columns in the dataset that were not directly related to our task of building a model for the total price of a airBnB over 4 nights. We also had to convert a number of the columns into workable data as many numeric values were stored as strings.
After building a workable dataset, the data was explored and visualizations were created to further develop an understanding of the data. We observed that there were dramatically different average prices depending on the neighbourhood. Interestingly, the properties in the city centre were not the most expensive as one might initially assume. Instead, **Huajiou**, which is located ~100km outside of the city centre, was the most expensive district. We realized this was due to it's proximity to the Great Wall, a highly popular tourist destination.
Additionally the property information was overlaid onto a map of Beijing to provide a visual aid to understand the distribution of AirBnb properties and their prices throughout Beijing.
In order to narrow down the variables to include in the regression models, correlation matrices were built. It was discovered there were 3 sets of variables that were highly intercorrelated. These were related to rating scores; number of reviews; and accommodation size. When we further examined these groups it was discovered the group related to accommodation size was significantly correlated with the price while the other two had very little correlation. It was also found that the correlations were even higher when the host is a superhost.
The regression model was developed through iteration and the final model's independent variables were **Max accommodation**, **neighbourhood**, **Number of Bedrooms**, **Room Type**, **Rating** which has a R^2^ value of **0.509** 
Our final regression model was used to predict the price for 2 people to rent for 4 nights from a property with a rating of greater than 90% and more than 10 reviews. The predicted price was ¥1,192 with a 95% confidence interval from ¥1,166 to ¥1,218.

```{r vroom_data, echo = FALSE, cache = TRUE,include=FALSE}
#Import the data
beijing_data <- vroom::vroom("http://data.insideairbnb.com/china/beijing/beijing/2020-06-19/data/listings.csv.gz")%>% 
  clean_names()
```


```{r missing_data_visual,include=FALSE}

# Select potentially relevant data columns (variables) to be further analyzed in the next steps
beijing_selected <- beijing_data %>% 
  #Select the relevant variables
  select(id, 
         host_response_time,
         host_response_rate,
         host_acceptance_rate,
         host_is_superhost,
         host_listings_count,
         host_total_listings_count,
         host_identity_verified,
         neighbourhood,
         neighbourhood_cleansed,
         zipcode,
         latitude,
         longitude,
         is_location_exact,
         property_type,
         room_type,
         accommodates,
         bathrooms,
         bedrooms,
         beds,
         price,
         security_deposit,
         cleaning_fee,
         guests_included,
         extra_people,
         minimum_nights,
         maximum_nights,
         number_of_reviews,
         number_of_reviews_ltm,
         review_scores_rating,
         review_scores_checkin,
         review_scores_cleanliness,
         review_scores_accuracy,
         review_scores_communication,
         review_scores_location,
         review_scores_value,
         instant_bookable,
         cancellation_policy,
         reviews_per_month) %>%
  # Perform basic mutate to change data type of numeric variables and parse number 
  mutate(price = parse_number(price),
         cleaning_fee = parse_number(cleaning_fee),
         security_deposit = parse_number(security_deposit),
         extra_people = parse_number(extra_people),
         host_response_rate = parse_number(host_response_rate),
         host_acceptance_rate = parse_number(host_acceptance_rate)
         )
```


```{r,include=FALSE}
beijing_cleaned <- beijing_selected %>%
  # Create a new variable prop_type_simplified
  mutate(prop_type_simplified = case_when(
    #The property type will be assigned to one of the top four if on the list,
    #or to Other if it isn't one of the top four
    property_type %in% c("Apartment","Condominium", "House","Loft") ~ property_type, 
    TRUE ~ "Other"),
    
    # Clean names of neighbourhoods ro be represented in latin letters
    neighbourhood_cleansed = case_when(
           neighbourhood_cleansed == "东城区" ~ "Dongcheng",
           neighbourhood_cleansed == "丰台区 / Fengtai" ~ "Fengtai",
           neighbourhood_cleansed == "大兴区 / Daxing" ~ "Daxing",
           neighbourhood_cleansed == "密云县 / Miyun" ~ "Miyun",
           neighbourhood_cleansed == "平谷区 / Pinggu" ~ "Pinggu",
           neighbourhood_cleansed == "延庆县 / Yanqing" ~ "Yanqing",
           neighbourhood_cleansed == "怀柔区 / Huairou" ~ "Huairou",
           neighbourhood_cleansed == "房山区" ~ "Fangshan",
           neighbourhood_cleansed == "昌平区" ~ "Changping",
           neighbourhood_cleansed == "朝阳区 / Chaoyang" ~ "Chaoyang",
           neighbourhood_cleansed == "海淀区" ~ "Haidian",
           neighbourhood_cleansed == "石景山区" ~ "Shijingshan",
           neighbourhood_cleansed == "西城区" ~ "Xicheng",
           neighbourhood_cleansed == "通州区 / Tongzhou" ~ "Tongzhou",
           neighbourhood_cleansed == "门头沟区 / Mentougou" ~ "Mentougou",
           neighbourhood_cleansed == "顺义区 / Shunyi" ~ "Shunyi")
    ) %>%
  #In the case we have NAs, give them the name N/A
  na_if("N/A")

#Assign 0s to the NA values for cleaning fee and security deposit cases
beijing_cleaned$cleaning_fee[is.na(beijing_cleaned$cleaning_fee)] <- 0
beijing_cleaned$security_deposit[is.na(beijing_cleaned$security_deposit)] <- 0

#Take a look at the cleaned dataset
skim(beijing_cleaned)
```



```{r,include=FALSE}
# refactor variables and filter for relevant AirBnBs

beijing_cleanbase <- beijing_cleaned %>%
  # We filter the dataset for listings where the minimum stay is lower or equal to 4 nights and where at least two people can be accommodated 
  filter(minimum_nights <= 4,
         accommodates >= 2,
         price != 0) %>% 
  # We perform multiple mutate operations in order to transform the variables into factor variables and relevel them
  mutate(host_response_time = fct_relevel(host_response_time,
                                            "within an hour", 
                                            "within a few hours",
                                            "within a day",
                                            "a few days or more"),
         cancellation_policy = fct_relevel(cancellation_policy,
                                           "flexible",
                                           "moderate",
                                           "strict_14_with_grace_period"),
         prop_type_simplified = fct_relevel(prop_type_simplified,
                                          "Apartment",
                                          "Condominium",
                                          "House",
                                          "Loft",
                                          "Other"),
         room_type = fct_relevel(room_type,
                                 "Shared room",
                                 "Private room",
                                 "Entire home/apt"),
         # We regroup the 16 neighbourhoods included in the neighbourhood_cleansed variable based on geographic location in Beijing, factoring in the median price per night for the particular district 
         neighbourhood_simplified = case_when(
           # no clear grouping possible for the following neighbourhoods, therefore name stays the same
           neighbourhood_cleansed == "Shunyi" ~ "Shunyi",
           neighbourhood_cleansed == "Chaoyang" ~ "Chaoyang",
           neighbourhood_cleansed == "Huairou" ~ "Huairou",
           # North east Beijing excl. Huairou due to significantly higher price point (Yanqing, Pinggu, Miyun) 
           neighbourhood_cleansed == "Yanqing" ~ "Northeast Beijing",
           neighbourhood_cleansed == "Pinggu" ~ "Northeast Beijing",
           neighbourhood_cleansed == "Miyun" ~ "Northeast Beijing",
           # Beijing central (Dongcheng, Xicheng)
           neighbourhood_cleansed == "Dongcheng" ~ "Central Beijing",
           neighbourhood_cleansed == "Xicheng" ~ "Central Beijing",
           # Western Beijing (Shijingshan, Haidian, Fengtai) 
           neighbourhood_cleansed == "Shijingshan" ~ "Western Beijing",
           neighbourhood_cleansed == "Haidian" ~ "Western Beijing",
           neighbourhood_cleansed == "Fengtai" ~ "Western Beijing",
           # Belt of Outskirts (Fangshan, Daxing, Tongzhou, Mentougou, Changping)
           neighbourhood_cleansed == "Mentougou" ~ "Beijing Outskirts",
           neighbourhood_cleansed == "Fangshan" ~ "Beijing Outskirts",
           neighbourhood_cleansed == "Changping" ~ "Beijing Outskirts",
           neighbourhood_cleansed == "Daxing" ~ "Beijing Outskirts",
           neighbourhood_cleansed == "Tongzhou" ~ "Beijing Outskirts"
           ),
         # set neighbourhood as factor
         neighbourhood_simplified = as.factor(neighbourhood_simplified),
         # Calculate the price for 4 nights for 2 people
         price_4_nights = case_when(guests_included >= 2 ~ (price*4+cleaning_fee),
                                    TRUE ~ ((price+extra_people)*4+cleaning_fee)),
         price_4_nights_log = log(price_4_nights),
         price_log = (log(price))
         ) %>%
  select(-neighbourhood, -property_type) %>% 
  filter(!is.na(host_is_superhost) | !is.na(host_identity_verified)) %>% 
  # We add an additional filter to remove all extreme outliers from the price, which we determined by adding 5x the interquartile range to the 3rd quartile
  filter(price < 10000)

skim(beijing_cleanbase)

```

# Regression Model

> We would like to know what would be the price for two people to stay at an Airbnb location for four nights. In order to make this prediction, we will run a linear regression with the price for four nights as the Y variable.

> Our goal is to come up with a model that explains the variability of the price for four nights. 
We will run various regressions in order to reach a model with the following characteristics :

> * The model has a high adjusted R-squared
* The model has statistically significant estimators
* The model's estimators are not collinear
* The model respects L-I-N-E assumptions of linear regression


> We first plot our variable price_4_nights in order to study its distribution.

```{r}

#We plot the distribution of price_4_nights
ggplot(beijing_cleanbase,aes(x=price_4_nights)) + geom_histogram() + theme_bw() + labs(title="The distribution of the price for 4 nights is right skewed", subtitle= "Distribution of the price for four nights", x="Price for four nights", y="Count")



```

> We notice that the distribution of price_4_nights is right skewed, indicating that we have a few outliers. Moreover, using price_4_nights could result in heteroskedasticity of the residuals. We could reduce the influence of the outliers and improve homoskedasticity by using a log scale, which is a more robust approach. Thus, we will use price_4_nights_log, which seems more symmetrical. 

```{r}

#We remove these 10 rows
beijing_cleanbase <- beijing_cleanbase %>%
  filter(price_4_nights!=0)
#We plot the distribution of price_4_nights_log
ggplot(beijing_cleanbase,aes(x=price_4_nights_log)) + geom_histogram() + theme_bw() + labs(title="The distribution of log(price_4_nights) is normally distributed", subtitle="Distribution of log(price_4_nights)",x="log(price_4_nights)", y="count")

```


> Let's try first to run the model with price_4_nights and to study the residuals

```{r}
#We try to run the model with price_4_nights, we check the residuals and the vif
model1 <- lm(price_4_nights ~ prop_type_simplified + number_of_reviews + review_scores_rating, data=beijing_cleanbase)
    msummary(model1) #We check the key numbers
    car::vif(model1) #We check the variance inflation factor
    autoplot(model1) + theme_bw() #We check the diagnostic plots


```

> This model has indeed many problems : 
* The linearity assumption is not respected, as there seems to be a pattern in the residuals vs fitted plot.
* Residuals do not follow a normal distribution 
* There are points with very high leverage that can influence our estimators.
* The variability of the random disturbance is different across the variable, which is a sign of heteroskedasticity. This indicates biases in our estimates.

> It wouldn't be relevant to interpret our coefficients in this model since they are most likely biased. Note that the model automatically used prop_type_apartment as the basis to avoid multicollinearity.

> As we said previously, we will run our regressions with price_4_nights_log.

```{r}

model2 <- lm(price_4_nights_log ~ prop_type_simplified + number_of_reviews + review_scores_rating, data=beijing_cleanbase)
msummary(model2)

car::vif(model2)
autoplot(model2) + theme_bw()
huxtable::huxreg(model1,model2)


```

> This model is slightly better:
* Residuals seem more randomly distributed
* The variance of residuals is constant
* Residuals follow a normal distribution 
* All independent variables are significant at the 0.1% level.
* The VIFs do not indicate multicollinearity (it was also the case for model 1 )

> Our independent variables are statistically significant (p-value<0.05) but our adjusted R-squared (0.082) is low, which is a sign of high variability within the model.However, we can still interpret our coefficients.

> We have to be careful with the interpretation since we used log(price_4_nights). We will have to use the exponential function to anti-log.

```{r}

#We anti-log to interpret the coefficients
condominium_vs_apart <- (exp(-0.062)-1)*100
condominium_vs_apart

house_vs_apart <- (exp(0.225)-1)*100
house_vs_apart

loft_vs_apart <- (exp(0.066)-1)*100
loft_vs_apart

other_vs_apart <- (exp(0.446)-1)*100
other_vs_apart

reviews <- (exp(-0.002)-1)*100
reviews

rating <- (exp(0.005)-1)*100
rating
```


> * We expect the price for 4 nights to be 6.01% cheaper for a condominium than for an apartment.
* We expect the price for 4 nights to be 25.2% more expensive for a house than for an apartment.
* We expect the price for 4 nights to be 6.82% more expensive for a loft than for an apartment.
* We expect the price for 4 nights to be -0.2% cheaper for other properties than for an apartment.
* We expect the price for 4 nights to be 0.501% more expensive for each additional rating point.


> Let us study if room_type is a significant predictor of the cost for 4 nights, given everything else in the model. We will fit a regression model (model3) that includes all the explanatory variables in model2 plus room_type.


```{r}

model3 <- lm(price_4_nights_log ~ prop_type_simplified + number_of_reviews + review_scores_rating + room_type, data=beijing_cleanbase)
car::vif(model3)
huxtable::huxreg(model2,model3)



```

> Note that the model automatically used room_typeSharedroom as the basis to avoid multicollinearity.
This model has a way higher adjusted R squared (0.261) than the previous one, but some estimators are not statistically significant (prop_type_simplifiedLoft). Also, the VIFs do not indicate any sign of multicollinearity.

> Room type is a significant predictor of the cost for 4 nights since p-value<0.05 for room_typePrivate room and for room_timeEntire home/apt. 

```{r}

#We anti-log to interpret the coefficients
roomprivate_vs_shared <- (exp(0.337)-1) *100
roomprivate_vs_shared

roomentire_vs_shared <- (exp(0.972)-1) *100
roomentire_vs_shared
```

> * We expect the price for 4 nights to be 40.1% more expensive for a private room than for a shared room.
* We expect the price for 4 nights to be 164% more expensive for an entire home than for a shared room.

> Let us now come up with more explanatory models that will hopefully have BLUE estimators.
> We will first try to see if bedrooms, bathrooms, beds and accommodates are good predictors of the price for 4 nights.

```{r}


model4 <- lm(price_4_nights_log ~ bedrooms + room_type, data=beijing_cleanbase)
model5 <- lm(price_4_nights_log ~ bedrooms + room_type+bathrooms , data=beijing_cleanbase)
model6 <- lm(price_4_nights_log ~ bedrooms+ room_type+ bathrooms+beds, data=beijing_cleanbase)
model7 <-lm(price_4_nights_log ~ bedrooms+ room_type +bathrooms+beds + accommodates, data=beijing_cleanbase)
  
huxtable::huxreg(model4,model5,model6,model7)
autoplot(model7) + theme_bw()
car::vif(model7)




```

> As we have seen in our correlation heatmap throughout the EDA, bedrooms, bathrooms and beds have an important correlation. It is however interesting to see that VIFs are not worrying since they are all below 5. Nevertheless, some of the chosen variables are not very good predictors of the price for 4 nights. 

> We first added bedrooms and room_type to our model and obtained an R squared of 0.283. We notice that the R squared does not increase significantly after adding bathrooms and beds, suggesting that adding only bedrooms in our final model would be the right choice. However, the variable accommodates seems to explain part of the variability in our dependent variable since the R squared increased to 0.369 after adding it, so we will also keep it. Our coefficients are all significant at the 0.1% level.

> This is our model with only bedrooms, room_type and accommodates:

```{r}

model8 <-lm(price_4_nights_log ~ bedrooms+ accommodates + room_type, data=beijing_cleanbase)
msummary(model8)
car::vif(model8)
autoplot(model8) + theme_bw()


```

> Our model with only bedrooms, room_type and accommodates is very simple and explains 36.4% of the variability of our dependent variable, and our estimators are significant at the 0.1% level. However, our estimates might be biased since there is a pattern in the residuals vs fitted plot, thus violating the linearity assumption.

> Let us now study if neighbourhoods add any value to our model after controlling for bedrooms, room_type and accommodates.

```{r}

model9 <-lm(price_4_nights_log ~ bedrooms+ room_type + accommodates+neighbourhood_simplified, data=beijing_cleanbase)
huxtable::huxreg(model8,model9)
car::vif(model9)
autoplot(model9) + theme_bw()

```

> The variable that is left out to avoid perfect multicollinearity is neighbourhood_Beijing_ outskirts.
The adjusted R-squared increases after adding neighbourhoods, which indicates that they improve our model. All our coefficients are significant at the 1% level, and the coefficients for neighbourhoods are all positive, indicating that all neighbourhoods are actually more expensive than Beijing Outskirts. There is still a concern with the linearity assumption and the influential points when looking at the distributions of residuals.

> Let us now see if superhosts command a pricing premium after controlling for other variables.

```{r}

model10 <- lm(price_4_nights_log ~ bedrooms+ accommodates + room_type  + neighbourhood_simplified + host_is_superhost, data=beijing_cleanbase)
huxtable::huxreg(model9,model10)
```

> Being a superhost has no effect on the price for 4 nights since the coefficient is not statistically significant at the 5% level. It is interesting to notice that the magnitude of the coefficient is very low and that the R squared does not increase at all from our previous model, indicating that this variable is not a good predictor of the price for 4 nights.


> Now, let us focus on reviews. It makes sense to only keep the total rating only since it is highly correlated with all the other review types (cleanliness, communication etc..)

```{r}

model11 <- lm(price_4_nights_log ~ accommodates + neighbourhood_simplified + bedrooms+ room_type + review_scores_rating  , data=beijing_cleanbase)

car::vif(model11)
huxtable::huxreg(model10,model11)

```

> Again, the R-squared increases drastically from 0.413 to 0.509 and our coefficients are all significant at the 0.1% level. The VIFs indicate that we have no collinearity so we will continue with these variables.

> We now add host response time to our model: 

```{r}

model12 <- lm(price_4_nights_log ~ accommodates + neighbourhood_simplified+ room_type   + bedrooms + review_scores_rating + host_response_time , data=beijing_cleanbase)
huxtable::huxreg(model11,model12)
```

> The R squared doesn't increase much and our estimators for "host_response_timewithin a day" and for "neighbourhood_simplifiedShunyi" are insignificant. Thus, we will be better off by dropping this variable.

Let us now add less intuitive variables and see if they actually influence the price for 4 nights.We will add : 

- Host_identity_verified : Is host's identity verified?
- is_location_exact : Is the location exact?
- instant_bookable : Is it instantly bookable?
- reviews_per_month : How many reviews per month does the property have?
- Security_deposit : What is the amount of the security deposit?
- Cancellation_policy : Is there a cancellation policy?
- Host_listings_count : How many properties has the host listed on Airbnb?

```{r}

model13 <- lm(price_4_nights_log ~ accommodates + neighbourhood_simplified + bedrooms+ room_type + review_scores_rating + host_identity_verified  + is_location_exact + instant_bookable + reviews_per_month + security_deposit + cancellation_policy +  host_listings_count, data=beijing_cleanbase)
huxtable::huxreg(model11,model13)
```

> All these variables only increase the R squared by 0.09, which is not significant at all. There is a tradeoff between having a simple model and having a precise model. The small increase in R squared does not justify the added complexity generated by these 7 new variables. Thus, we will decide to leave them out of our model.

> For now, our best model has been model11.

> It explains 50.9% of the variability in the price for 4 nights. The VIFs are all below 5, indicating that there is no sign of multicollinearity. All our coefficients are significant at the 0.1% level.

> However, since we are great statisticians, we must also criticize our model. We have tried to correct for outliers in our dataset but some may still be troublesome as they have an impact on the assumptions of linear regression. Some points have a high leverage, residuals do not follow perfectly a normal distribution and there are some minors sign of heteroskedasticity. All in all, we believe that our model has a high explanatory power but that some of our coefficients may be slightly biased. 

```{r}


autoplot(model11) + theme_bw()


```

> This is our final summary table:

```{r}
#We rearrange models for the final table
model_d= lm(price_4_nights_log ~ accommodates + neighbourhood_simplified + bedrooms+ room_type, data=beijing_cleanbase)
model_c= lm(price_4_nights_log ~ accommodates + neighbourhood_simplified + bedrooms, data=beijing_cleanbase)
model_b= lm(price_4_nights_log ~ accommodates + neighbourhood_simplified, data=beijing_cleanbase)
model_a= lm(price_4_nights_log ~ accommodates, data=beijing_cleanbase)

huxtable::huxreg(model_a,model_b,model_c,model_d,model11)

car::vif(model11)
```


## Price prediction 

> As described in the previous section, we choose model 11 for the price prediction for our particular search as it is the most robust model while being comparably simple in terms of number of indendent variables.  

```{r airbnb_target}

# Filter Dataset for criteria

airbnb_target_search <- beijing_cleanbase %>% 
  filter(room_type == "Private room",
         number_of_reviews >= 10,
         review_scores_rating >= 90)

# Apply regression model for the different 
# Calculate the predicted price and the corresponding 95% CIs

predicted_price_4_nights <- predict(model11, newdata = airbnb_target_search, interval = "confidence") %>% 
  # antilog the values
  exp() %>% 
  data.frame() %>%
  summarize(lower_bound = mean(lwr),
            predicted_price = mean(fit),
            upper_bound = mean(upr))
  
predicted_price_4_nights %>%
  kbl(col.names = c("Lower Bound (¥)", "Predicted Price (¥)", "Upper Bound (¥)"), format.args = list(big.mark = ",")) %>%
  kable_material(c("striped", "hover")) %>%
  kable_styling(fixed_thead = T)

```

> Based on running the preferred model11 for predicting the price for 4 nights in an AirBnB in Beijing with the given criteria (reviews >= 10, average rating >= 90 and private room), we calculated the 95% CI based on the filtered dataset accordingly. This results in an expected price of ¥1,192 (Lower bound = ¥1,166, Upper bound = ¥1,218).
